Properties of Winning Iterated Prisoner's Dilemma Strategies.

### 2020-04-25 paper/main.tex
minor copy edits to introduction

Conceptualising strategies and understanding the best way of playing the game
has been of interest to the scientific community since the formulation of the
game in 1950~\cite{Flood1958}. Following the computer tournaments of R. Axelrod in the
1980's~\cite{Axelrod1980a, Axelrod1980b}, a strategy's performance in a round
robin computer tournament became a common evaluation technique for newly designed
strategies. Many tournaments have followed Axelrod's~\cite{Beaufils1997, Bendor1991,
Harper2017, Kendall2007, Stephens2002, Stewart2012} and today more hundreds of strategies exist
in the literature.

@@

Other work has built upon the limitations of TFT, and in some cases have shown
that suggestions made by R. Axelrod did not necessarily apply in alternative environmental settings.
In~\cite{Bendor1991, Donninger1986, Molander1985, Hammerstein1984} it was shown
that TFT suffered in environments with noise. This was
mainly due to the strategy's lack of generosity and contrition. Since TFT immediately
punishes a defection, in a noisy environment it can get stuck in a
repeated cycle of defections and cooperations. Some new strategies, more
robust in tournaments with noise, were soon introduced, including
Nice and Forgiving~\cite{Bendor1991}, Generous Tit For Tat~\cite{Nowak1992},
and Pavlov (aka Win Stay Lose Shift)~\cite{Nowak1993}, as well as later
variants such as OmegaTFT \cite{kendall2007iterated}. The introduction of a new strategy
is often accompanied by a claim that the new strategy is the best known, despite
only being tested against a small number of opponents or against specific classes of
opponents not necessarily representative of all possible or all published strategies.
The lack of testing against formally defined strategies and tournament winners
is understandable given the effort required to implement the hundreds of published IPD strategies
Implementing prior strategies faithfully is often extremely difficult or
impossible due to insufficient descriptions and lack of published
implementations or code.
This calls into question any claims of superiority or robustness of newly introduced strategies.

A set of envious IPD strategies were introduced called zero-determinant strategies (ZDs)
in~\cite{Press2012}. These strategies attempt to force a linear relationship between
stationary payoffs against other memory-one opponents, potentally ensuring that they
receive a higher average payout. While ZDs were introduced with a small tournament in
which some were reportedly successful \cite{Stewart2012}, this result has not generally
held in future work. In~\cite{Harper2017} a series of strategies trained using
reinforcement learning were introduced, and a tournament containing over 200 strategies
featured no ZDs ranking in top spots. Instead, the top ranked strategies
were a set of ``clever'' (in the sense of R. Axelrod's characteristics) trained
strategies based on lookup tables~\cite{Axelrod1987}, hidden Markov
models~\cite{Harper2017}, and finite state automata~\cite{Miller1996}.
Similarly, in \cite{mathieu2017}, a set of evolutionarily-trained strategies,
and a pre-selected set of known strategies, outperformed a collection of 12 ZDs.

@@

Compared to other works, where typically a few selected or introduced strategies
are evaluated on a small number of tournaments and/or small number of opponents,
this manuscript evaluates the performance of \numberofstrategies
strategies in \numberofalltournaments tournaments. Furthermore a large portion
of the strategies used in this manuscript are drawn from known and named strategies
in IPD literature, including many previous tournament winners,
in contrast to other work that may have randomly generated many essentially arbitrary
strategies (typically restrained to a class such as memory-one strategies,
or those of a certain structural form such as finite state machines or deterministic
memory two strategies). Additionally, our tournaments come in a
number of variations including standard tournaments emulating R. Axelrod's original tournaments,
tournaments with noise, probabilistic match length, and both noise and probabilistic match length.
This diversity of strategies and tournament types yields new insights and tests
earlier claims in alternative settings against known powerful strategies.

>>>>

Conceptualising strategies and understanding the best way of playing the game
has been of interest to the scientific community since the formulation of the
game in 1950~\cite{Flood1958}. Following the computer tournaments of R. Axelrod in the
1980's~\cite{Axelrod1980a, Axelrod1980b}, a strategy's performance in a round
robin computer tournament became a common evaluation technique for newly designed
strategies. Many tournaments have followed Axelrod's~\cite{Beaufils1997, Bendor1991,
Harper2017, Kendall2007, Stephens2002, Stewart2012} and today hundreds of more strategies exist
in the literature.

@@

Other work has built upon the limitations of TFT, and in some cases have shown
that suggestions made by R. Axelrod did not necessarily apply in alternative environmental settings.
In~\cite{Bendor1991, Donninger1986, Molander1985, Hammerstein1984} it was shown
that TFT suffered in environments with noise. This was
mainly due to the strategy's lack of generosity and contrition. Since TFT immediately
punishes a defection, in a noisy environment it can get stuck in a
repeated cycle of defections and cooperations. Some new strategies, more
robust in tournaments with noise, were soon introduced, including
Nice and Forgiving~\cite{Bendor1991}, Generous Tit For Tat~\cite{Nowak1992},
and Pavlov (aka Win Stay Lose Shift)~\cite{Nowak1993}, as well as later
variants such as OmegaTFT \cite{kendall2007iterated}. The introduction of a new strategy
is often accompanied by a claim that the new strategy is the best known, despite
only being tested against a small number of opponents or against specific classes of
opponents not necessarily representative of all possible or all published strategies.
The lack of testing against formally defined strategies and tournament winners
is understandable given the effort required to implement the hundreds of published IPD strategies.
Implementing prior strategies faithfully is often extremely difficult or
impossible due to insufficient descriptions and lack of published
implementations or code.
This calls into question any claims of superiority or robustness of newly introduced strategies.

A set of envious IPD strategies were introduced called zero-determinant strategies (ZDs)
in~\cite{Press2012}. These strategies attempt to force a linear relationship between
stationary payoffs against other memory-one opponents, potentially ensuring that they
receive a higher average payout. While ZDs were introduced with a small tournament in
which some were reportedly successful \cite{Stewart2012}, this result has not generally
held in future work. In~\cite{Harper2017} a series of strategies trained using
reinforcement learning were introduced, and a tournament containing over 200 strategies
featured no ZDs ranking in top spots. Instead, the top ranked strategies
were a set of ``clever'' (in the sense of R. Axelrod's characteristics) trained
strategies based on lookup tables~\cite{Axelrod1987}, hidden Markov
models~\cite{Harper2017}, and finite state automata~\cite{Miller1996}.
Similarly, in \cite{mathieu2017}, a set of evolutionarily-trained strategies,
and a pre-selected set of known strategies, outperformed a collection of 12 ZDs.

@@

Compared to other works, where typically a few selected or introduced strategies
are evaluated on a small number of tournaments and/or small number of opponents,
this manuscript evaluates the performance of \numberofstrategies
strategies in \numberofalltournaments tournaments. Furthermore, a large portion
of the strategies used in this manuscript are drawn from known and named strategies
in IPD literature, including many previous tournament winners,
in contrast to other work that may have randomly generated many essentially arbitrary
strategies (typically restrained to a class such as memory-one strategies,
or those of a certain structural form such as finite state machines or deterministic
memory-two strategies). Additionally, our tournaments come in a
number of variations including standard tournaments emulating R. Axelrod's original tournaments,
tournaments with noise, probabilistic match length, and both noise and probabilistic match length.
This diversity of strategies and tournament types yields new insights and tests
earlier claims in alternative settings against known powerful strategies.
